---
title: "Benthic Habitats"
author: "Florian Mayer"
date: "06/12/2015"
output: html_document
---

# Preparation

This workbook lives inside the MPA report source code. 
It is under [version control]() and also uploads itself to the 
[internal data catalogue]().

To prepare the environment:

* the work dir is set to the respective folder inside the MPA reports (the convention is to 
 keep the MPA reports underneath `~/projects`),
* source custom functions (eg. loading a CSV from CKAN),
* source the secret CKAN settings (CKAN URL and API key),
* load required libraries.

```{r}
setwd("~/projects/data-pipelines/scripts/indicators/benthic_habitats")
source("~/projects/data-pipelines/setup/ckan.R")
source("~/projects/data-pipelines/setup/ckan_setup.R")
library(tidyr) # clean data
library(dplyr) # munge clean data
library(ggplot2) # visualise munged data
library(DT) # DataTables widget
library(vegan) # Multivariate community data analysis
```

Next, the CKAN resource IDs of the raw data CSV, the workbook, and the output PDF
figure are defined.

```{r}
csv_rid <- "216cdeb3-7c94-46bd-be62-531e2b469c57"
pdf_rid = "40ca93b6-09e5-457a-bb8a-3e884b61d75d"
pdf_fn <- "habitat.pdf"
txt_rid = "efbe602f-8800-4198-be63-eed01e03338f"
txt_fn <- "benthic_habitats.Rmd"
```

# Load data
Now, the raw data are read from CKAN and parsed into correct data types.

```{r}
d <- load_ckan_csv(csv_rid, date_colnames = c('date', 'Date', 'time', 'Time'))

dplyr::tbl_df(d)
dplyr::glimpse(d)
```

# Create site profiles
Data is grouped by Surveys (individual samplings of transects = "Replicate").
We'll retain the spatial hierarchy of a surveyed transect's Park, Sector, Zone 
and Site, as well as the temporal coordinates (Date).

With the survey code (which contains the full spatial hierarchy as well as the Date)
as `row.names`, data are split up into site profiles as responding variables `X` 
(abundance of each taxon) and factor variables for each site `Y` in absence of 
real explanatory variables.

```{r}
sp3 <- d %>% 
  tbl_df() %>% 
  group_by(Survey, Date, Park, Sector, Zone, Site, ReplicateCode, Level3Class) %>%
  tally() %>%
  spread(Level3Class, n, fill=0)

row.names(sp3) <- sp3$Survey

Y = select(sp3, -Survey, -Date, -Park, -Sector, -Zone, -Site, -ReplicateCode)
X = select(sp3, Date, Park, Sector, Zone, Site, ReplicateCode)
```

Now `X` and `Y` look like this:

```{r, echo=FALSE}
datatable(X)
datatable(Y)
```

# Ordination
Let's run a non-metric multidimensional scaling (NMDS) for demo purposes 
(don't interpret the result).

```{r}
Y.mds <- metaMDS(Y)
Y.mds
```

Let's plot the results and save the figure as a PDF.

```{r}
plot(Y.mds, type="n", main="Non-metric multidimensional scaling")
points(Y.mds, display="sites", cex=0.8, pch=21, col="red", bg="yellow")
text(Y.mds, display="spec", cex=0.7, col="blue")

pdf(pdf_fn, height = 7, width = 7)
plot(Y.mds, type="n", main="Non-metric multidimensional scaling")
points(Y.mds, display="sites", cex=0.8, pch=21, col="red", bg="yellow")
text(Y.mds, display="spec", cex=0.7, col="blue")
dev.off()
```

# Post analysis: Upload and cleanup

* Create a file `ckan_secret` with this line containing your `CKAN API_KEY`:
* ```library(ckanr); ckanr::ckanr_setup(url="http://internal-data.dpaw.wa.gov.au/", key="API_KEY")```
* The PDF and R script must already exist on CKAN (if not, create them once)
* Upload the figure (PDF) and this script (TXT) to CKAN
* Set workdir to main report location

```{r}
# ckanr::resource_update(pdf_rid, pdf_fn)
# ckanr::resource_update(txt_rid, txt_fn)
setwd("~/projects/data-pipelines")
```
