---
title: "Accessing BioSys data"
author: "Florian Mayer, DBCA"
date: "2017/09/21"
output: html_document
---

This workbook shows how to access EcoPAAS data once they are uploaded
into the data warehouse BioSys.

The two primary purposes of doing so are:

* to QA data and identify possible required edits to records, and
* to analyse and visualise data.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs, echo=FALSE, message=FALSE}
library(httr)
library(dplyr)
library(purrr)
library(tibble)
library(DT)
# devtools::install_github("datadotworld/datapackage-r")
# library(datapkg)
# install.packages("kimisc")
library(kimisc)
```
# Setup
Obtain a Biosys username and password and add to your `~/.Rprofile`:

```{r setenv, eval=FALSE}
Sys.setenv(BIOSYS_API_USERNAME = "USERNAME")
Sys.setenv(BIOSYS_API_PASSWORD = "PASSWORD")
```

# Data flow

* Benthic images are analysed and annotated in the software EcoPAAS
* EcoPAAS outputs data as Excel
* BioSys imports Excel using a config created by Paul's uploader (tm)
* [Biosys GUI](https://biosys.dbca.wa.gov.au/)
* [BioSys API in Browser](https://biosys.dbca.wa.gov.au/sso/api) 
using single sign on (needs active browser window)
* [BioSys API in scripts](https://biosys.dbca.wa.gov.au/api) using basicauth 
(username and password)
* [BioSys API documentation](https://biosys-admin.dbca.wa.gov.au/api/explorer/)

# Accessing data from BioSys
## Helper functions

* Data is retrieved from the BioSys API through an HTTP GET with basicauth.
* The JSON returned from the BioSys API is parsed into a `tibble`.
* All heavy lifting is factored out into helper functions.

```{r helpers}
#' Make map_chr resilient against NULL
map_chr_hack <- function(.x, .f, ...) {
  map(.x, .f, ...) %>%
    purrr::map_if(is.null, ~ NA_character_) %>%
    purrr::flatten_chr()
}


#' Make map_dbl resilient against NULL
map_dbl_hack <- function(.x, .f, ...) {
  map(.x, .f, ...) %>%
    purrr::map_if(is.null, ~ NA_real_) %>%
    purrr::flatten_dbl()
}


#' Basic Biosys API GET returning HTTP response content
biosys_get <- function(api_path,
  api="https://biosys.dbca.wa.gov.au/api/",
  auth=httr::authenticate(Sys.getenv("BIOSYS_UN"), 
                          Sys.getenv("BIOSYS_PW")),
  query = list()){
  httr::GET(file.path(api, api_path), auth, query = query) %>% 
    httr::content(.)
} 


#' Parse BioSys projects to tibble
biosys_projects <- function(
  auth=httr::authenticate(
    Sys.getenv("BIOSYS_UN"), Sys.getenv("BIOSYS_PW"))){
  biosys_get("projects", auth = auth) %>% {
    tibble::tibble(
      id = purrr::map_chr(., "id"),
      name = purrr::map_chr(., "name"),
      code = purrr::map_chr(., "code"),
      description = map_chr_hack(., "description"),
      site_count = purrr::map_int(., "site_count"),
      dataset_count = purrr::map_int(., "dataset_count"),
      record_count = purrr::map_int(., "record_count"),
      longitude = purrr::map(., c("centroid", "coordinates")) %>% 
        map_dbl_hack(magrittr::extract2, 1),
      latitude = purrr::map(., c("centroid", "coordinates")) %>% 
        map_dbl_hack(magrittr::extract2, 2),
      datum = purrr::map_chr(., "datum"),
      timezone = purrr::map_chr(., "timezone")
      # site_data_package
      # custodians
    )
  }
}


#' Parse BioSys datasets to tibble
biosys_datasets <- function(
  auth=httr::authenticate(
    Sys.getenv("BIOSYS_UN"), Sys.getenv("BIOSYS_PW")),
  project_id = NULL){
  if (!is.null(project_id)) {
    query = list("project" = project_id)
  } else {
      query = list()
  }
  biosys_get("datasets", auth = auth, query = query) %>% {
    tibble::tibble(
      id = purrr::map_chr(., "id"),
      record_count = purrr::map_int(., "record_count"),
      data_package = purrr::map(., "data_package"),
      name = purrr::map_chr(., "name"),
      type = purrr::map_chr(., "type"),
      description = map_chr_hack(., "description"),
      project_id = purrr::map_int(., "project")
    )
  }
}


#' Parse BioSys records to tibble
biosys_records <- function(
  auth=httr::authenticate(Sys.getenv("BIOSYS_UN"),
                          Sys.getenv("BIOSYS_PW")),
  project_id = NULL){
  if (!is.null(project_id)) {
    query = list("dataset__project__id" = project_id)
  } else {
      query = list()
  }
  biosys_json <- biosys_get("records", auth = auth, query = query)
  
  biosys_records <- biosys_json %>% {
   tibble::tibble(
     id = purrr::map_chr(., "id"),
     datetime = map_chr_hack(., "datetime"),
     species_name = map_chr_hack(., "species_name"),
     name_id = map_chr_hack(., "name_id"),
     file_name = purrr::map(., "source_info") %>% 
        map_chr_hack(magrittr::extract2, 1),
     file_row = purrr::map(., "source_info") %>% 
       map_chr_hack(magrittr::extract2, 2),
     last_modified = purrr::map_chr(., "last_modified"),
     dataset = purrr::map_chr(., "dataset"),
     site = map_chr_hack(., "site"),
     data = purrr::map(., "data")
   )}
  
  metadatacols <- biosys_records %>% select(-data)
  
  datacols <- biosys_records$data %>% 
    kimisc::list_to_df(.) %>% 
    magrittr::extract2("value") %>% 
    purrr::transpose(.) %>% 
    tibble::as_tibble(.)
  
  out <- cbind(metadatacols, datacols)
  out
}


#' Parse BioSys records into a tibble, flatten data into cols
parse_records <- function(biosys_json){
 biosys_records <- biosys_json %>% {
   tibble::tibble(
     id = purrr::map_chr(., "id"),
     datetime = map_chr_hack(., "datetime"),
     species_name = map_chr_hack(., "species_name"),
     name_id = map_chr_hack(., "name_id"),
     file_name = purrr::map(., "source_info") %>% 
        map_chr_hack(magrittr::extract2, 1),
     file_row = purrr::map(., "source_info") %>% 
       map_chr_hack(magrittr::extract2, 2),
     last_modified = purrr::map_chr(., "last_modified"),
     dataset = purrr::map_chr(., "dataset"),
     site = map_chr_hack(., "site"),
     data = purrr::map(., "data")
   )}
  
  datacols <- biosys_records$data %>% 
  kimisc::list_to_df(.) %>% 
  magrittr::extract2("value") %>% 
  purrr::transpose(.) %>% 
  tibble::as_tibble(.)
  
  metadatacols <- biosys_records %>% select(-data)

  cbind(metadatacols, datacols)
}
```

## BioSys projects
```{r projects}
projects <- biosys_projects()
DT::datatable(projects)
```

## BioSys datasets
```{r datasets}
datasets <- biosys_datasets(project_id = 6)
DT::datatable(datasets)
```

## BioSys records
```{r records}
records <- biosys_records(project_id = 6)
DT::datatable(head(records, n=100))
```

